\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-05-17 redaction/dhp_sept2016.tex
changement

In this section, we investigate the role of the additive component in total transport costs by exploring the time dimension of our data. As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}). Two main comments can be made. First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974. Second, the decrease appears more important for ocean than air transportation: Transportation costs were divided by 3 for the ocean shipping, versus by 2 for air transport. That said, it may be the case the oil shock inflated transportation costs that specific year. If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for air transport costs, and 2-3 percentage points for vessel transport costs (depending whether we consider mean or median estimates). These results are consistent with \citet{hummels2007}.

Before making any definite statement about this, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i°)} the evolution of per product and per partner transport costs and \textit{ii°)} the evolution of the composition of trade. Total transport costs may thus have decreased over time because the share of neighbour countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii°)}), independently of any change in transport costs \textit{per se}. As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e. per product- and per partner- transport costs.

>>>>

In this section, we investigate the role of the additive component in total transport costs in a historical perspective. As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}). Two main comments can be made. First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974. \todo{Guillaume, ameliorer les choses} If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for air transport costs, and 2-3 percentage points for vessel transport costs (depending whether we consider mean or median estimates). These results are consistent with \citet{hummels2007}.

Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i°)} the evolution of per product and per partner transport costs and \textit{ii°)} the evolution of the composition of trade. Total transport costs may thus have decreased over time because the share of neighbour countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii°)}), independently of any change in transport costs \textit{per se}. As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``pure'' international transport costs, i.e. per product- and per partner- transport costs.
